---
Title: EDA Analysis on an advertising dataset
Author: Joy Kiriba
Date: 02/07/2021
Output: html_document
editor_options:
  chunk_output_type: inline
---

#1. Defining the Question

##a.Specifying the data analytic question
``` Identify which individuals are most likely to click on ads.```

##b) Defining the metric of success
``` When we are able to identify the individuals likely to click on ads based on the information and data given.```

##c) Understanding the context
```A Kenyan entrepreneur has created an online cryptography course and would want to advertise it on her blog. She currently targets audiences originating from various countries. In the past, she ran ads to advertise a related course on the same blog and collected data in the process. She would now like to employ my services as a Data Science Consultant to help her identify which individuals are most likely to click on her ads.```

##d) Recording the experimental design
```The process will entail: 
* Define the question, the metric for success, the context, experimental design taken.

* Read and explore the given dataset.

* Define the appropriateness of the available data to answer the given question.

* Find and deal with outliers, anomalies, and missing data within the dataset.

* Perform univariate and bivariate analysis recording my observations.```

##e) Data Relevance

```The data provided is very relevant and has a wide range of variables to consider in achieving our main objective of this project. 

[Advertising dataset](http://bit.ly/IPAdvertisingData)```


#2. Reading the Data
```{r}

# Loading our data set
advertising <- read.csv('advertising.csv')

```

#3. Checking the data
```{r}
# Viewing the top 6 entries
head(advertising)

#viewing the whole data set
View(advertising)

#Data types of the columns
str(advertising)

#Statistical summary of the data set
summary(advertising)

#checking the number of entries and attributes
dim(advertising)

#checking the class of our object
class(advertising)

```
#4. Tidying the data/ Data Cleaning

```*Checking for missing data*```
```{r}
#Checking for missing data
colSums(is.na(advertising))

```
```There are no missing values in our data set.```

```*Checking for duplicates*```
```{r}
duplicated.rows <- advertising[duplicated(advertising),]
duplicated.rows

```
```There is no duplicated data in our dataset```

```*Checking for outliers*```

```{r}
#Create a list of numeric columns
num.cols <- list(advertising$Daily.Time.Spent.on.Site,advertising$Age,
advertising$Area.Income,advertising$Daily.Internet.Usage,advertising$Male, advertising$Clicked.on.Ad)

#Checking for outliers
boxplot(num.cols, names=c('Daily.Time.Spent.on.Site', 'Age', 'Area.Income', 'Daily.Internet.Usage', 'Male', 'Clicked.on.Ad'), main='Boxplots to show Outliers', las=2)

#Listing the outliers
boxplot.stats(advertising$Area.Income)$out

```
```{r}
#Plotting boxplots of individual columns

boxplot(advertising$Daily.Time.Spent.on.Site, main='Boxplot of Daily time spent on site', xlab='Daily Time spent on the site', ylab='value')
boxplot(advertising$Age, main='Boxplot of age', xlab='Age', ylab='Value')
boxplot(advertising$Area.Income, main='Boxplot of area income', xlab='Area income', ylab='Value')
boxplot(advertising$Daily.Internet.Usage, main='Boxplot of Daily Internet Usage', xlab='Daily Internet Usage', ylab='Value')
boxplot(advertising$Male, main='Boxplot of Gender', xlab='Gender', ylab='Value')
boxplot(advertising$Clicked.on.Ad, main='Boxplot of clicked on ad', xlab='Clicked on ad', ylab='Value')
 
```
```The outliers in area income might be due to low numbers of add clicks so no need to remove them.```

#5 Univariate Exploratory Data Analysis
## Measures of Central Tendancy

```{r}
#Finding the mean
colMeans(advertising[sapply(advertising, is.numeric)])

#Finding the median
#loading the tidyverse and robustbase(for the colMedians function) libraries
library(robustbase)
library(tidyverse)
advertising%>%
  select_if(is.numeric) %>%
  as.matrix()%>%
  colMedians()

#Finding the mode
mode <- function(x) {
  uniq_data <- unique(x)
  map_data <- match(x, uniq_data)
  tab_data <- tabulate(map_data)
  max_val <- max(tab_data)
  uniq_data[tab_data == max_val]
}

mode(advertising$Daily.Time.Spent.on.Site)
mode(advertising$Age)
mode(advertising$Area.Income)
mode(advertising$Daily.Internet.Usage)
mode(advertising$Ad.Topic.Line)
mode(advertising$City)
mode(advertising$Male)
mode(advertising$Country)
mode(advertising$Timestamp)
mode(advertising$Clicked.on.Ad)

```
## Measures of Dispersion
```{r}
#Finding the minimum
num.cols <- list(advertising$Daily.Time.Spent.on.Site,advertising$Age,
advertising$Area.Income,advertising$Daily.Internet.Usage,advertising$Male, advertising$Clicked.on.Ad)

min(advertising$Daily.Time.Spent.on.Site)
min(advertising$Age)
min(advertising$Area.Income)
min(advertising$Daily.Internet.Usage)
min(advertising$Male)
min(advertising$Clicked.on.Ad)

#Finding the maximum
max(advertising$Daily.Time.Spent.on.Site)
max(advertising$Age)
max(advertising$Area.Income)
max(advertising$Daily.Internet.Usage)
max(advertising$Male)
max(advertising$Clicked.on.Ad)

#Finding the Range
range(advertising$Daily.Time.Spent.on.Site)
range(advertising$Age)
range(advertising$Area.Income)
range(advertising$Daily.Internet.Usage)
range(advertising$Male)
range(advertising$Clicked.on.Ad)

#Finding the quantiles
quantile(advertising$Daily.Time.Spent.on.Site)
quantile(advertising$Age)
quantile(advertising$Area.Income)
quantile(advertising$Daily.Internet.Usage)
quantile(advertising$Male)
quantile(advertising$Clicked.on.Ad)

#Finding the variance
var(advertising$Daily.Time.Spent.on.Site)
var(advertising$Age)
var(advertising$Area.Income)
var(advertising$Daily.Internet.Usage)
var(advertising$Male)
var(advertising$Clicked.on.Ad)

#Finding the Standard Deviation
sd(advertising$Daily.Time.Spent.on.Site)
sd(advertising$Age)
sd(advertising$Area.Income)
sd(advertising$Daily.Internet.Usage)
sd(advertising$Male)
sd(advertising$Clicked.on.Ad)

```
## Univariate Data analysis Graphicals
```{r}
#Creating a frequency table for city
city <- advertising$City
freq_city <- table(city)
freq_city

#Creating a frequency table for country
country <- advertising$Country
freq_country <- table(country)
freq_country

#Creating a histogram for daily time spent on site
hist(advertising$Daily.Time.Spent.on.Site)

#Creating a histogram for age
hist(advertising$Age)

#Creating a histogram for Area Income
hist(advertising$Area.Income)

#Creating a histogram of daily internet usage
hist(advertising$Daily.Internet.Usage)

#Creating a histogram for Male
hist(advertising$Male)

#Frequency table of Male
table(advertising$Male)

#Creating a histogram for clicked on ad
hist(advertising$Clicked.on.Ad)

#Frequency table of clicked on ad
table(advertising$Clicked.on.Ad)


```
### Observations from Univariate Analysis
```
* Most people in the dataset are between the age of 25 and 40 years.
* The highest frequency of area income is between 60000 and 70000.
* Males are more than female but not with a big margin.
* Those who clicked on the ad are equal to those that did not.
* Highest daily internet usage is 269.96.
* The most frequent time spent on site is 80

```
#6. Bivariate and Multivariate Data Analysis
```{r}
#Finding the covariance
cov <- cov(advertising[, unlist(lapply(advertising, is.numeric))])
round(cov, 3)

#Finding the correlation
cor <- cor(advertising[, unlist(lapply(advertising, is.numeric))])
round(cor, 3)


```
## Bivariate and Multivariate Graphicals
```{r}
#Plotting a scatterplot of daily time spent on site and clicked on ad
#num.cols <- list(advertising$Daily.Time.Spent.on.Site,advertising$Age,
#advertising$Area.Income,advertising$Daily.Internet.Usage,advertising$Male, advertising$Clicked.on.Ad)
daily.time.spent.on.site <- advertising$Daily.Time.Spent.on.Site
clicked.on.ad <- advertising$Clicked.on.Ad

plot(daily.time.spent.on.site, clicked.on.ad, xlab='Daily time spent on site', ylab='Clicked on ad')

#Plotting a scatter plot of male against clicked on ad
male <- advertising$Male
clicked.on.ad <- advertising$Clicked.on.Ad
plot(male, clicked.on.ad, xlab='male', ylab='clicked on ad')

#Plotting area income and clicked on add
area.income <- advertising$Area.Income
clicked.on.ad <- advertising$Clicked.on.Ad
plot(area.income, clicked.on.ad, xlab='Area income', ylab='Clicked on ad')

#Plotting scatter plot of age and clicked on ad
age <- advertising$Age
clicked.on.ad <- advertising$Clicked.on.Ad
plot(age, clicked.on.ad, xlab='Age', ylab='Clicked on label')

#Plottin scatter plot of daily internet usage and clicked on ad
daily.internet.usage <- advertising$Daily.Internet.Usage
clicked.on.ad <- advertising$Clicked.on.Ad
plot(daily.internet.usage, clicked.on.ad, xlab='Daily internet usage', ylab='Clicked on ad')

```

```{r}
#Plotting a heat map using the correlation matrix
heatmap(x=cor, symm=TRUE)

#Plotting a correlogram
library('corrplot')
corrplot(cor, type='upper', order='hclust', tl.color='black', tl.srt=45)
```

### Observations from Bivariate and Multivariate Analysis
```
* Daily internet usage and daily time spent on site are positively correlated.

* Daily time spent on site and Daily Internet usage are negatively correlated to whether an individual clicks on an ad or not.

* There is no strong correlation between male(gender), age, area income and whether an individual clicked on an ad or not.
```
## Conclusion
```
* The age and gender do not determine whether an individual clicks on an ad. This is probably because their interests on the internet are different from what the ad is about.

* Daily time spent on a site has a negative correlation on whether an individual clicks on an ad probably because they are already on the site and are aware of what the ad is about.
```
## Recommendations
```
* I would recommend further research to be done to add more variables to the data set that would help determine whether an individual is more likely to click on an ad. This is because the present variables do not help us answer that question.

* A classification model could be used to predict the likelihood of an individual clicking on an ad, thus giving more insight.
```
